Asymmetric tail-association in extreme climatic events affects population-extinction risk

### 2019-10-12 MainText.Rmd
rearranging figs...continued

We found density-dependent metapopulation models in stochastic 
environment showed similar patterns for extinction risk as density-independent 
metapopulation model showed in recent study [@Ghosh_copula]. For each density-dependent 
model listed in Table \ref{tab_model_summary}, we observed higher extinction risk for 
the metapopulations when spatially correlated noises were left-tail dependent than the one with right-tail association.
But that observation remained valid when the model parameter - here considered as growth rate $(r)$ - was 
sufficiently low so that under-compensatory dynamics (monotonous approach towards
its stable equilibrium) for the deterministic analog of the model was ensured.
We observed right tail dependent noise caused higher extinction 
risk for the metapopulations than left tail dependent noise when the model 
was in over-compensatory regime (oscillatory approach towards stable equilibrium) for higher 
value of $r$. Figs. \ref{fig_extrisk_allmodel_localdisp_etd} and \ref{SM-fig_extrisk_allmodel_localdisp_mtd} show
variation of extinction risk against local dispersal rate ($d$) for six density dependent metapopulation models 
(see Table \ref{tab_model_summary}) with growth rate $r$ considered either from under-compensatory or 
from over-compensatory regime. We also got similar results with our prediction (see Q2 from \nameref{Introduction})
for global dispersal (Figs. \ref{SM-fig_extrisk_allmodel_globaldisp_etd} and \ref{SM-fig_extrisk_allmodel_globaldisp_mtd}). 
Figs. (\ref{SM-fig_ricker_risk_local_disp_scl_1_etd} - \ref{SM-fig_risk_local_disp_abrewer_r_vary_scl_0.6_mtd}) 
demonstrate how extinction risk changed against local dispersal rate for a whole range
of $r$, whereas Figs. (\ref{SM-fig_ricker_risk_global_disp_scl_1_etd} - 
\ref{SM-fig_risk_global_disp_abrewer_r_vary_scl_0.6_mtd}) 
represent the same but for global dispersal among patches. 
Both case showed the signature of "rescue-effect" (i.e. reduced extinction risk for 
finite dispersal compared with $d=0$).

@@

 <!--FIG : age structure model results : linear, nonlinear-->
\begin{figure*}[!h]
\textbf{ \hspace{5.4 cm} (A) \hspace{4 cm} (B)} \\
\vspace{-0.9 cm}
\begin{center}
\includegraphics[width=5 cm]{./Results/age_str_results/extrisk_age_str_extremetail_linear.pdf}
\includegraphics[width=5 cm]{./Results/age_str_results/extrisk_age_str_extremetail_nonlinear_omg_100.pdf}\\
\caption{Extinction risk against time from single patch age structured $(A)$ linear ($\Omega = \infty$) and $(B)$ nonlinear ($\Omega = 100$) models for $10000$ simulations; shading represents 95\% confidence interval. Fecundity of adults and survival of eggs have either extreme left or extreme right tail association. \label{fig_extrisk_agestr_etd}}
\end{center}
\end{figure*}

<!--FIG :  summary : all nonlinear model results for local dispersion : extrisk vs. d-->
\begin{figure*}[!h]
\begin{center}
\includegraphics[width=15 cm]{./Results/allmodel_ext_risk_local_disp_extremenoise.pdf}
\caption{Extinction risks are computed after $25$ years for six nonlinear density-dependence models (Table \ref{tab_model_summary}) with local dispersal among $n=5$ patches and for $10000$ simulations. Extinction risk varies as dispersion-rate ($d$) goes from $0$ to $0.5$. Last two columns represent extinction risk calculated for intrinsic growth rate $(r)$ taken from under-compensatory and over-compensatory regime of deterministic model analog, respectively.\label{fig_extrisk_allmodel_localdisp_etd}}
\end{center}
\end{figure*}

 <!--FIG : all nonlinear model results with extreme noise, d=0, extrisk vs. r-->
\begin{figure*}[!h]
\begin{center}
\includegraphics[width=10 cm]{./Results/common_legend_extreme_noise.pdf}\\
\includegraphics[width=5 cm]{./Results/ricker_scl_1/ricker_extrisk_d0_extremenoise.pdf}
\includegraphics[width=5 cm]{./Results/hassell_scl_1/hassell_extrisk_d0_extremenoise.pdf}\\
\vspace{-0.6 cm}
\includegraphics[width=5 cm]{./Results/msmith_scl_0.8/msmith_extrisk_d0_extremenoise.pdf}
\includegraphics[width=5 cm]{./Results/pennycuick_scl_1/pennycuick_extrisk_d0_extremenoise.pdf}\\
\vspace{-0.6 cm}
\includegraphics[width=5 cm]{./Results/verhulst_scl_0.6/verhulst_extrisk_d0_extremenoise.pdf}
\includegraphics[width=5 cm]{./Results/austinbrewer_scl_0.6/abrewer_extrisk_d0_extremenoise.pdf}\\
\caption{Transition of higher to lower extinction risk (calculated after $25$ years) for 
the metapopulations affected by extreme left tail dependent noise across
$5$ multiple patches without dispersal ($d=0$), as intrinsic 
growth rate ($r$) goes from under to over-compensatory regime. Extinction risk for 
extreme right tail dependent noise follows the opposite trend.
Error bars represent the standard errors associated with $10000$
simulations for extinction risk computed at each $r$. Dotted line represents the value of $r$
for each model where their deterministic analog switches from under- to over-compensatory dynamics.
\label{fig_extrisk_d0_etd}}
\end{center}
\end{figure*}

>>>>

We found density-dependent metapopulation models in stochastic 
environment showed similar patterns for extinction risk as density-independent 
metapopulation model showed in recent study [@Ghosh_copula]. For each density-dependent 
model listed in Table \ref{tab_model_summary}, we observed higher extinction risk for 
the metapopulations when spatially correlated noises were left-tail dependent than the one with right-tail association.
But that observation remained valid when the model parameter - here considered as growth rate $(r)$ - was 
sufficiently low so that under-compensatory dynamics (monotonous approach towards
its stable equilibrium) for the deterministic analog of the model was ensured.
We observed right tail dependent noise caused higher extinction 
risk for the metapopulations than left tail dependent noise when the model 
was in over-compensatory regime (oscillatory approach towards stable equilibrium) for higher 
value of $r$. Figs. \ref{fig_extrisk_allmodel_localdisp} shows
variation of extinction risk against local dispersal rate ($d$) for six density dependent metapopulation models 
(see Table \ref{tab_model_summary}) with growth rate $r$ considered either from under-compensatory or 
from over-compensatory regime. We also got similar results with our prediction (see Q2 from \nameref{Introduction})
for global dispersal (Fig. \ref{SM-fig_extrisk_allmodel_globaldisp}). 
Figs. (\ref{SM-fig_ricker_risk_local_disp_scl_1_etd} - \ref{SM-fig_risk_local_disp_abrewer_r_vary_scl_0.6_mtd}) 
demonstrate how extinction risk changed against local dispersal rate for a whole range
of $r$, whereas Figs. (\ref{SM-fig_ricker_risk_global_disp_scl_1_etd} - 
\ref{SM-fig_risk_global_disp_abrewer_r_vary_scl_0.6_mtd}) 
represent the same but for global dispersal among patches. 
Both case showed the signature of "rescue-effect" (i.e. reduced extinction risk for 
finite dispersal compared with $d=0$).

@@

 <!--FIG : age structure model results : linear, nonlinear-->
\begin{figure*}[!h]
\textbf{ \hspace{5.4 cm} (A) \hspace{4 cm} (B)} \\
\vspace{-0.9 cm}
\begin{center}
\includegraphics[width=5 cm]{./Results/age_str_results/extrisk_age_str_extremetail_linear.pdf}
\includegraphics[width=5 cm]{./Results/age_str_results/extrisk_age_str_extremetail_nonlinear_omg_100.pdf}\\
\end{center}
\textbf{ \hspace{5.4 cm} (C) \hspace{4 cm} (D)} \\
\vspace{-0.9 cm}
\begin{center}
\includegraphics[width=5 cm]{./Results/age_str_results/extrisk_age_str_C_SC_linear.pdf}
\includegraphics[width=5 cm]{./Results/age_str_results/extrisk_age_str_C_SC_nonlinear_omg_100.pdf}
\caption{Extinction risk against time from single patch age structured linear ($\Omega = \infty$) $(A, C)$ and nonlinear ($\Omega = 100$) $(C, D)$ models for $10000$ simulations; shading represents 95\% confidence interval. $(A, B)$ show extinction risk with time when fecundity of adults and survival of eggs have extreme left tail (LT) or extreme right tail (RT) association, $(C, D)$ show the analogous scenario but with moderate tail association between vital rates. \label{fig_extrisk_agestr}}
\end{center}
\end{figure*}

<!--FIG :  summary : all nonlinear model results for local dispersion : extrisk vs. d-->
\begin{figure*}[!h]
\begin{center}
\includegraphics[width=14.5 cm]{./Results/allmodel_ext_risk_local_disp_extreme_and_moderate_noise.pdf}
\caption{Extinction risks are computed after $25$ years for six nonlinear density-dependence models (Table \ref{tab_model_summary}) with local dispersal among $n=5$ patches and for $10000$ simulations. Spatially correlated 
noises have varying degree of tail associations (extreme left tail, moderate left tail, moderate right tail and extreme right tail) as introduced in Fig. \ref{fig_intro_cop}. Last two columns represent variation of extinction risk against dispersal rate ($d$), calculated for intrinsic growth rate $(r)$ taken from under-compensatory and over-compensatory regime of deterministic model analog.\label{fig_extrisk_allmodel_localdisp}}
\end{center}
\end{figure*}

<!--FIG : all nonlinear model results, d=0, extrisk vs. r-->
\begin{figure*}[!h]
\begin{center}
\includegraphics[width=12 cm]{./Results/common_legend_extreme_and_moderate_noise.pdf}\\
\includegraphics[width=6 cm]{./Results/ricker_scl_1/ricker_extrisk_d0_extreme_and_moderate_noise.pdf}
\includegraphics[width=6 cm]{./Results/hassell_scl_1/hassell_extrisk_d0_extreme_and_moderate_noise.pdf}\\
\vspace{-0.6 cm}
\includegraphics[width=6 cm]{./Results/msmith_scl_0.8/msmith_extrisk_d0_extreme_and_moderate_noise.pdf}
\includegraphics[width=6 cm]{./Results/pennycuick_scl_1/pennycuick_extrisk_d0_extreme_and_moderate_noise.pdf}\\
\vspace{-0.6 cm}
\includegraphics[width=6 cm]{./Results/verhulst_scl_0.6/verhulst_extrisk_d0_extreme_and_moderate_noise.pdf}
\includegraphics[width=6 cm]{./Results/austinbrewer_scl_0.6/abrewer_extrisk_d0_extreme_and_moderate_noise.pdf}\\
\caption{Transition of higher to lower extinction risk (calculated after $25$ years) for 
the metapopulations affected by extreme left tail dependent noise across
$5$ multiple patches without dispersal ($d=0$), as intrinsic 
growth rate ($r$) goes from under to over-compensatory regime. Extinction risk for 
extreme right tail dependent noise follows the opposite trend.
Error bars represent the standard errors associated with $10000$
simulations for extinction risk computed at each $r$. Dotted line represents the value of $r$
for each model where their deterministic analog switches from under- to over-compensatory dynamics.
\label{fig_extrisk_d0_etd}}
\end{center}
\end{figure*}
